---
title: Detection of extreme heatwaves and low sea ice cover in the Antarctic Peninsula  for
  evaluation of effects over Krill
author: "Lucas Krüger, Maurício Mardones, Lorena Rebolledo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#---------- Download NOAA erddap optimum interpolation SST----

#this part of the script is a modification of the scripts in the heatwaveR vignette
#available at https://robwschlegel.github.io/heatwaveR/articles/OISST_preparation.html


```{r, echo =FALSE}
#---------- Download NOAA erddap optimum interpolation SST----

# The packages we will use
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(patchwork)
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing
library(raster)
library(terra)
library(sf)

library(mapdata)
library(mapproj)

```

# The information for the NOAA OISST data

```{r, echo=FALSE}


rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

rerddap::info(datasetid = "erdMH1chlamday", url = "https://coastwatch.pfeg.noaa.gov/erddap/")



```

# Download and prepare data based on user provided start and end dates using three different requests per geographical strata so it can be run on slow PCs.

```{r}


# spatial units to use


shp<-shapefile("D:/HeatWaves/Krill/Data/Shapefiles/Clipped_StrataGCS.shp")


crs=CRS( "+proj=longlat +datum=WGS84 +no_defs") # coordinate reference system CRS

shp2<-spTransform(shp,crs) # be sure it is in the correct coordinate system


#load antarctic peninsula shapefile to plot it
# this shapefile was cut from high resolution file from the SCAR Antarctic Digital Database, 2023
# Gerrish, L., Ireland, L., Fretwell, P., & Cooper, P. (2023). High resolution vector polygons of the Antarctic coastline (7.7) [Data set]. UK Polar Data Centre, Natural Environment Research Council, UK Research & Innovation. https://doi.org/10.5285/0be5339c-9d35-44c9-a10f-da4b5356840b'

wap<-shapefile("D:/HeatWaves/Krill/Data/Shapefiles/peninsula_correct.shp") 


# plot map with shapefile (Fig 1)
shpdf<-data.frame(shp2)

ggplot()+
  
  xlab("Longitude")+ylab("Latitude")+
  theme_bw()+
    geom_polygon(data=shp2,aes(x = long, y = lat, group = group,label=group),fill="lightblue",colour="black")+
  geom_polygon(data=wap,aes(x = long, y = lat, group = group))+
geom_text(data=shpdf,aes(label=ID, 
            x=Labx,
            y=Laby),colour="red2",
            size = 3, hjust=0, vjust=-1) +
      coord_quickmap(xlim=c(-75,-50),ylim=c(-75, -59), expand = TRUE, clip = "on")


shp.sf<-st_as_sf(shp)
unique(shp.sf$ID)

#Extract coordinates and generates a bounding box 

geom<-st_geometry(shp.sf)

bbox_list <- lapply(st_geometry(shp.sf), st_bbox)

#min and max coordinates of each strata as data frame

df1<-data.frame(t(as.data.frame((bbox_list[[1]])))) # Elephant
df2<-data.frame(t(as.data.frame((bbox_list[[2]])))) # South Shetlands
df3<-data.frame(t(as.data.frame((bbox_list[[3]])))) # Bransfield
df4<-data.frame(t(as.data.frame((bbox_list[[4]])))) #Joinville
df5<-data.frame(t(as.data.frame((bbox_list[[5]])))) # Outer 
df9<-data.frame(t(as.data.frame((bbox_list[[9]])))) #Extra SW

```

#download and process data for Elephant Island Strata

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df1$ymin),max(df1$ymax)), 
                                longitude = c(min(df1$xmin),max(df1$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_EL1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_EL2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_EL3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_EL1, file = "D:/HeatWaves/Krill/Data/OISST_EL1.Rds")
base::saveRDS(OISST_EL2, file = "D:/HeatWaves/Krill/Data/OISST_EL2.Rds")
base::saveRDS(OISST_EL3, file = "D:/HeatWaves/Krill/Data/OISST_EL3.Rds")

### summarize data per day

EL1<-plyr::ddply(OISST_EL1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EL2<-plyr::ddply(OISST_EL2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EL3<-plyr::ddply(OISST_EL3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

EL12<-rbind(EL1,EL2)
EL123<-rbind(EL12,EL3)

base::saveRDS(EL123, file = "D:/HeatWaves/Krill/Data/Elephant_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_EL1","OISST_EL2","OISST_EL3","OISST_sub_dl","shp","maxmin","bbox_list",
          "EL1","EL2","EL3","EL12","EL123"))

gc()



```
#download and process data for South Shetland West Strata

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df2$ymin),max(df2$ymax)),
                                longitude = c(min(df2$xmin),max(df2$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_SS1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_SS2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_SS3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_SS1, file = "D:/HeatWaves/Krill/Data/OISST_SS1.Rds")
base::saveRDS(OISST_SS2, file = "D:/HeatWaves/Krill/Data/OISST_SS2.Rds")
base::saveRDS(OISST_SS3, file = "D:/HeatWaves/Krill/Data/OISST_SS3.Rds")

### summarize data per day

SS1<-plyr::ddply(OISST_SS1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SS2<-plyr::ddply(OISST_SS2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SS3<-plyr::ddply(OISST_SS3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

SS12<-rbind(SS1,SS2)
SS123<-rbind(SS12,SS3)

base::saveRDS(SS123, file = "D:/HeatWaves/Krill/Data/West_SSI_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_SS1","OISST_SS2","OISST_SS3","OISST_sub_dl",
          "SS1","SS2","SS3","SS12","SS123"))

gc()



```


#download and process data for Bransfield Strait Strata

```{r, echo=FALSE}

df3$ID[3]

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df3$ymin),max(df3$ymax)),
                                longitude = c(min(df3$xmin),max(df3$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_BS1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_BS2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_BS3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_BS1, file = "D:/HeatWaves/Krill/Data/OISST_BS1.Rds")
base::saveRDS(OISST_BS2, file = "D:/HeatWaves/Krill/Data/OISST_BS2.Rds")
base::saveRDS(OISST_BS3, file = "D:/HeatWaves/Krill/Data/OISST_BS3.Rds")


### summarize data per day

BS1<-plyr::ddply(OISST_BS1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

BS2<-plyr::ddply(OISST_BS2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

BS3<-plyr::ddply(OISST_BS3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

BS12<-rbind(BS1,BS2)
BS123<-rbind(BS12,BS3)

base::saveRDS(BS123, file = "D:/HeatWaves/Krill/Data/Bransfield_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_BS1","OISST_BS2","OISST_BS3","OISST_sub_dl",
          "BS1","BS2","BS3","BS12","BS123"))

gc()



```




#download and process data for Joinville Island Strata

```{r, echo=FALSE}



OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df4$ymin),max(df4$ymax)),
                                longitude = c(min(df4$xmin),max(df4$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_JI1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_JI2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_JI3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_JI1, file = "D:/HeatWaves/Krill/Data/OISST_JI1.Rds")
base::saveRDS(OISST_JI2, file = "D:/HeatWaves/Krill/Data/OISST_JI2.Rds")
base::saveRDS(OISST_JI3, file = "D:/HeatWaves/Krill/Data/OISST_JI3.Rds")


OISST_JI1<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI1.Rds")
OISST_JI2<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI2.Rds")
OISST_JI3<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI3.Rds")


### summarize data per day

JI1<-plyr::ddply(OISST_JI1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

JI2<-plyr::ddply(OISST_JI2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

JI3<-plyr::ddply(OISST_JI3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

JI12<-rbind(JI1,JI2)
JI123<-rbind(JI12,JI3)

base::saveRDS(JI123, file = "D:/HeatWaves/Krill/Data/Joinvile_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_JI1","OISST_JI2","OISST_JI3","OISST_sub_dl",
          "JI1","JI2","JI3","JI12","JI123"))

gc()



```


#download and process data for Extra Strata (Gerlache?)

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df5$ymin),max(df5$ymax)),
                                longitude = c(min(df5$xmin),max(df5$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_EX1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_EX2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_EX3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_EX1, file = "D:/HeatWaves/Krill/Data/OISST_EX1.Rds")
base::saveRDS(OISST_EX2, file = "D:/HeatWaves/Krill/Data/OISST_EX2.Rds")
base::saveRDS(OISST_EX3, file = "D:/HeatWaves/Krill/Data/OISST_EX3.Rds")

### summarize data per day

EX1<-plyr::ddply(OISST_EX1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EX2<-plyr::ddply(OISST_EX2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EX3<-plyr::ddply(OISST_EX3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

EX12<-rbind(EX1,EX2)
EX123<-rbind(EX12,EX3)

base::saveRDS(EX123, file = "D:/HeatWaves/Krill/Data/Extra_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_EX1","OISST_EX2","OISST_EX3","OISST_sub_dl",
          "EX1","EX2","EX3","EX12","EX123"))

gc()



```


#download and process data for Outer SW Strata (Gerlache?)

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df9$ymin),max(df9$ymax)),
                                longitude = c(min(df9$xmin),max(df9$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_SW1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_SW2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_SW3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_SW1, file = "D:/HeatWaves/Krill/Data/OISST_SW1.Rds")
base::saveRDS(OISST_SW2, file = "D:/HeatWaves/Krill/Data/OISST_SW2.Rds")
base::saveRDS(OISST_SW3, file = "D:/HeatWaves/Krill/Data/OISST_SW3.Rds")

### summarize data per day

SW1<-plyr::ddply(OISST_SW1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SW2<-plyr::ddply(OISST_SW2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SW3<-plyr::ddply(OISST_SW3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

SW12<-rbind(SW1,SW2)
SW123<-rbind(SW12,SW3)

base::saveRDS(SW123, file = "D:/HeatWaves/Krill/Data/OuterSW_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_SW1","OISST_SW2","OISST_SW3","OISST_sub_dl",
          "SW1","SW2","SW3","SW12","SW123"))

gc()



```


###----------- calculating marine heatwaves-------------


```{r,echo=FALSE}

library(heatwaveR)


#load and merge data

BS<-base::readRDS("D:/HeatWaves/Krill/Data/Bransfield_Strata.Rds")
EI<-base::readRDS("D:/HeatWaves/Krill/Data/Elephant_Strata.Rds")
EX<-base::readRDS("D:/HeatWaves/Krill/Data/Extra_Strata.Rds")
JI<-base::readRDS("D:/HeatWaves/Krill/Data/Joinvile_Strata.Rds")
SS<-base::readRDS("D:/HeatWaves/Krill/Data/West_SSI_Strata.Rds")
SW<-base::readRDS("D:/HeatWaves/Krill/Data/OuterSW_Strata.Rds")

### calculate thresholds from climatology

# tMax thresholds
tMax_clim.BS <- ts2clm(data = BS, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)

tMax_clim.EI <- ts2clm(data = EI, 
                     y = tMax, climatologyPeriod = c("1986-01-01", "2022-12-31"), pctile = 90)

tMax_clim.EX <- ts2clm(data = EX, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)

tMax_clim.JI <- ts2clm(data = JI, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)

tMax_clim.SS <- ts2clm(data = SS, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)

tMax_clim.SW <- ts2clm(data = SW, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)



# iMin thresholds

sicMean_clim.BS <- ts2clm(data = BS, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 20)
sicMean_clim.EI <- ts2clm(data = EI, 
                     y = sicMean, climatologyPeriod = c("1986-01-01", "2022-12-31"), pctile = 20)
sicMean_clim.EX <- ts2clm(data = EX, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 20)

sicMean_clim.JI <- ts2clm(data = JI, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 20)
sicMean_clim.SS <- ts2clm(data = SS, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 20)
sicMean_clim.SW <- ts2clm(data = SW, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 20)



#DETECT alldfS
eBS.t <- detect_event(data = tMax_clim.BS, y = tMax,minDuration = 5,maxGap = 7) 
eEI.t <- detect_event(data = tMax_clim.EI, y = tMax,minDuration = 5,maxGap = 7) 
eEX.t <- detect_event(data = tMax_clim.EX, y = tMax,minDuration = 5,maxGap = 7) 
eJI.t <- detect_event(data = tMax_clim.JI, y = tMax,minDuration = 5,maxGap = 7) 
eSS.t <- detect_event(data = tMax_clim.SS, y = tMax,minDuration = 5,maxGap = 7) 
eSW.t <- detect_event(data = tMax_clim.SW, y = tMax,minDuration = 5,maxGap = 7) 


eBS.sic <- detect_event(data = sicMean_clim.BS, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eEI.sic <- detect_event(data = sicMean_clim.EI, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eEX.sic <- detect_event(data = sicMean_clim.EX, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eJI.sic <- detect_event(data = sicMean_clim.JI, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eSS.sic <- detect_event(data = sicMean_clim.SS, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eSW.sic <- detect_event(data = sicMean_clim.SW, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 


#export climatologies and create dataframes

bsclimT<-data.frame(eBS.t$climatology,sector=c("BS")) 
elclimT<-data.frame(eEI.t$climatology,sector=c("EI")) 
exclimT<-data.frame(eEX.t$climatology,sector=c("EX")) 
jiclimT<-data.frame(eJI.t$climatology,sector=c("JI")) 
ssclimT<-data.frame(eSS.t$climatology,sector=c("SS")) 
swclimT<-data.frame(eSW.t$climatology,sector=c("SW")) 

bsclimS<-data.frame(eBS.sic$climatology,sector=c("BS")) 
elclimS<-data.frame(eEI.sic$climatology,sector=c("EI")) 
exclimS<-data.frame(eEX.sic$climatology,sector=c("EX")) 
jiclimS<-data.frame(eJI.sic$climatology,sector=c("JI")) 
ssclimS<-data.frame(eSS.sic$climatology,sector=c("SS")) 
swclimS<-data.frame(eSW.sic$climatology,sector=c("SW")) 

climT<-rbind(bsclimT,elclimT,exclimT,jiclimT,ssclimT,swclimT) # join data
climS<-rbind(bsclimS,elclimS,exclimS,jiclimS,ssclimS,swclimS) # join data

# plots for temperature and sea ice

# Create category breaks and select slice of data.frame


# Set line colours
lineColCat <- c(
  "Temperature" = "black",
  "Climatology" = "gray20",
  "Threshold" = "darkgreen",
  "2x Threshold" = "darkgreen",
  "3x Threshold" = "darkgreen",
  "4x Threshold" = "darkgreen"
  )

# Set category fill colours
fillColCat <- c(
  "Moderate" = "#ffc866",
  "Strong" = "#ff6900",
  "Severe" = "#9e0000",
  "Extreme" = "#2d0000"
  )

# Set category fill colours
fillColCat2 <- c(
  "Moderate" = "#ffc866",
  "Normal" = "#85B7CC"
  )

lineColCat2 <- c(
  "Sea Ice" = "black",
  "Climatology" = "blue3",
  "Threshold" = "red3"
  )
# Create category breaks 
climT
climT$diff = climT$thresh - climT$seas
climT$thresh_2x = climT$thresh + climT$diff
climT$thresh_3x = climT$thresh_2x + climT$diff
climT$thresh_4x = climT$thresh_3x + climT$diff


climS
climS$diff = climS$thresh - climS$seas
climS$thresh_2x = climS$thresh + climS$diff
climS$thresh_3x = climS$thresh_2x + climS$diff
climS$thresh_4x = climS$thresh_3x + climS$diff


saveRDS(climS,"D:/HeatWaves/Krill/Data/climatologyICE.Rds")
saveRDS(climT,"D:/HeatWaves/Krill/Data/climatologyTEMP.Rds")


#figure 2
(ggplot(data = subset(climT,t>'2022-03-31'), aes(x = t, y = tMax)) +
  geom_flame(aes(y2 = thresh, fill = "Moderate")) +
  geom_flame(aes(y2 = thresh_2x, fill = "Strong")) +
  geom_flame(aes(y2 = thresh_3x, fill = "Severe")) +
  geom_flame(aes(y2 = thresh_4x, fill = "Extreme")) +
  geom_line(aes(y = thresh_2x, col = "2x Threshold"), size = 0.7, linetype = "dashed") +
  geom_line(aes(y = thresh_3x, col = "3x Threshold"), size = 0.7, linetype = "dotdash") +
  geom_line(aes(y = thresh_4x, col = "4x Threshold"), size = 0.7, linetype = "dotted") +
  geom_line(aes(y = seas, col = "Climatology"), size = 0.7) +
  geom_line(aes(y = thresh, col = "Threshold"), size = 0.7) +
  geom_line(aes(y = tMax, col = "Temperature"), size = 0.6) +
  scale_colour_manual(name = NULL, values = lineColCat,
                      breaks = c("Temperature", "Climatology", "Threshold",
                                 "2x Threshold", "3x Threshold", "4x Threshold")) +
  scale_fill_manual(name = NULL, values = fillColCat, guide = FALSE) +
  scale_x_date(date_labels = "%b %Y") +
  guides(colour = guide_legend(override.aes = list(linetype = c("solid", "solid", "solid",
                                                                "dashed", "dotdash", "dotted"),
                                                   size = c(0.6, 0.7, 0.7, 0.7, 0.7, 0.7)))) +
  labs(y = "Temperature [°C]", x = NULL)+
  facet_wrap(sector~.)+theme_bw()+theme(panel.spacing.x = unit(2, "lines"))+
  ggtitle(label="a. High temperature extremes"))/
  
  
  
  
  (ggplot(data = subset(climS,t>'2022-03-31'), aes(x = t, y = sicMean)) +
     geom_flame(aes(y = sicMean,y2=thresh, fill = "Normal")) +
     geom_flame(aes(y = thresh,y2=sicMean, fill = "Moderate")) +
     geom_line(aes(y = seas, col = "Climatology"), size = 0.7,linetype="dashed") +
     geom_line(aes(y = thresh, col = "Threshold"), size = 0.7,linetype="dotdash") +
     geom_line(aes(y = sicMean, col = "Sea Ice"), size = 0.6) +
     scale_colour_manual(name = NULL, values = lineColCat2,
                         breaks = c("Sea Ice", "Climatology","Threshold")) +
     scale_fill_manual(name = NULL, values = fillColCat2, guide = FALSE) +
     scale_x_date(date_labels = "%b %Y") +
     guides(colour = guide_legend(override.aes = list(linetype = c("solid","dashed",
                                                                   "dotdash")))) +
     labs(y = "Sea ice cover [fraction]", x = NULL)+
     facet_wrap(sector~.)+theme_bw()+theme(panel.spacing.x = unit(2, "lines"))+
     
     ggtitle(label="b. Low sea ice extremes"))

#export events and creat data frame

bsevenT<-data.frame(eBS.t$event,sector=c("BS")) 
elevenT<-data.frame(eEI.t$event,sector=c("EI")) 
exevenT<-data.frame(eEX.t$event,sector=c("EX")) 
jievenT<-data.frame(eJI.t$event,sector=c("JI")) 
ssevenT<-data.frame(eSS.t$event,sector=c("SS")) 
swevenT<-data.frame(eSW.t$event,sector=c("SW")) 

bsevenS<-data.frame(eBS.sic$event,sector=c("BS")) 
elevenS<-data.frame(eEI.sic$event,sector=c("EI")) 
exevenS<-data.frame(eEX.sic$event,sector=c("EX")) 
jievenS<-data.frame(eJI.sic$event,sector=c("JI")) 
ssevenS<-data.frame(eSS.sic$event,sector=c("SS")) 
swevenS<-data.frame(eSW.sic$event,sector=c("SW")) 

evenT<-rbind(bsevenT,elevenT,exevenT,jievenT,ssevenT,swevenT) # join events data
evenS<-rbind(bsevenS,elevenS,exevenS,jievenS,ssevenS,swevenS) # join events data

evenT$Day<-(decimal_date(evenT$date_peak)-year(evenT$date_peak))*365 # day of the year
evenT$quart<-quarter(evenT$date_peak) # tri-mensual classification

evenS$Day<-(decimal_date(evenS$date_peak)-year(evenS$date_peak))*365 # day of the year
evenS$quart<-quarter(evenS$date_peak) # tri-mensual classification


# change names
evenT$Season[evenT$quart=="1"]<-"JFM"
evenT$Season[evenT$quart=="2"]<-"AMJ"
evenT$Season[evenT$quart=="3"]<-"JAS"
evenT$Season[evenT$quart=="4"]<-"OND"

evenS$Season[evenS$quart=="1"]<-"JFM"
evenS$Season[evenS$quart=="2"]<-"AMJ"
evenS$Season[evenS$quart=="3"]<-"JAS"
evenS$Season[evenS$quart=="4"]<-"OND"

evenT$Season<-factor(evenT$Season,levels=c("JFM","AMJ","JAS","OND")) # order levels

evenS$Season<-factor(evenS$Season,levels=c("JFM","AMJ","JAS","OND")) # order levels

#PLOT figure 3



(ggplot(subset(evenT,date_peak>'2009-12-31'),aes(date_peak,duration,colour=Season,linetype=Season))+
    geom_lolli()+
    theme_bw()+ylab("Duration (days)")+
    scale_colour_manual(values=c("red","green4","blue","orange"))+
    facet_wrap(sector~.)+
    xlab("Date peak")+
    ggtitle(label="a. Duration of high temperature events"))+
  
  (ggplot(subset(evenT,date_peak>'2009-12-31'),aes(date_peak,intensity_max,colour=Season,linetype=Season))+
     geom_lolli()+
     theme_bw()+ylab("Intensity (°C)")+
     scale_colour_manual(values=c("red","green4","blue","orange"))+
     facet_wrap(sector~.)+
     xlab("Date peak")+
     ggtitle(label="b. Intensity of temperature events"))+
  
  (ggplot(subset(evenS,date_peak>'2009-12-31'),aes(date_peak,duration,colour=Season,linetype=Season))+
     geom_lolli()+
     theme_bw()+ylab("Duration (days)")+
     scale_colour_manual(values=c("red","green4","blue","orange"))+
     facet_wrap(sector~.)+
     xlab("Date peak")+
     ggtitle(label="c. Duration of low sea ice events"))+


  (ggplot(subset(evenS,date_peak>'2009-12-31'),aes(date_peak,intensity_max,colour=Season,linetype=Season))+
     geom_lolli()+
     theme_bw()+ylab("Intensity")+
     scale_colour_manual(values=c("red","green4","blue","orange"))+
     facet_wrap(sector~.)+
     xlab("Date peak")+
     ggtitle(label="d. Intensity of low sea ice events"))



head(evenT)
evenT$Year<-year(evenT$date_start)
evenS$Year<-year(evenS$date_start)



#summarize per season and year

evenT.S<-plyr::ddply(evenT, c("Year","Season","sector"), summarise,
            TN=length(duration),
            TmeanDur=mean(duration),
            TcumDur=sum(duration),
            TmeanInt=mean(intensity_mean),
            TmaxInt=max(intensity_max))

evenS.S<-plyr::ddply(evenS, c("Year","Season","sector"), summarise,
            SN=length(duration),
            SmeanDur=mean(duration),
            ScumDur=sum(duration),
            SmeanInt=mean(intensity_mean),
            SmaxInt=max(intensity_max))


evesum<-merge(evenT.S,evenS.S,by=c("Year","Season","sector"))

saveRDS(evesum,"D:/HeatWaves/Krill/Data/events_summary.Rds")

summary(evesum)

evesum$TN[is.na(evesum$TN)]<-0
evesum$TmeanDur[is.na(evesum$TmeanDur)]<-0
evesum$TcumDur[is.na(evesum$TcumDur)]<-0
evesum$TmeanInt[is.na(evesum$TmeanInt)]<-0
evesum$TmaxInt[is.na(evesum$TmaxInt)]<-0

evesum$SN[is.na(evesum$SN)]<-0
evesum$SmeanDur[is.na(evesum$SmeanDur)]<-0
evesum$ScumDur[is.na(evesum$ScumDur)]<-0
evesum$SmeanInt[is.na(evesum$SmeanInt)]<-0
evesum$SmaxInt[is.na(evesum$SmaxInt)]<-0


summary(evesum)

ggplot(data=evesum)+
    geom_smooth(data=evesum,aes(x=log(TmeanDur+1),y=log(SmeanDur+1)),
              method="gam",se=F)+
  

  geom_point(data=evesum,aes(x=log(TmeanDur+1),y=log(SmeanDur+1),colour=Season,shape=Season))+
  facet_wrap(sector~.)+theme_bw()+

ggplot(data=evesum)+
    geom_smooth(data=evesum,aes(x=TmaxInt,y=SmaxInt),
              method="gam",se=F)+
 

  geom_point(data=evesum,aes(x=TmaxInt,y=SmaxInt,colour=Season,shape=Season))+
  facet_wrap(sector~.)+theme_bw()

### there is no  correspondence between temperature and sea ice in most sectors (delayed response? see fig 2)

head(BS)

BS$sector<-c("BS")
EI$sector<-c("EI")
EX$sector<-c("EX")
JI$sector<-c("JI")
SS$sector<-c("SS")
SW$sector<-c("SW")


alldf<-rbind(BS,EI,EX,JI,SS,SW)

alldf$month<-month(alldf$t)
alldf$Year<-year(alldf$t)
alldf$quarter<-quarter(alldf$t)

alldf$Season[alldf$quart=="1"]<-"JFM"
alldf$Season[alldf$quart=="2"]<-"AMJ"
alldf$Season[alldf$quart=="3"]<-"JAS"
alldf$Season[alldf$quart=="4"]<-"OND"


adfm<-plyr::ddply(alldf, c("Year","Season","sector"), summarise,
            tMax=mean(tMax),
            sicMin=mean(sicMin))


#figure 4
ggplot(adfm,aes(tMax,sicMin))+
  geom_point(alpha=0.05)+ 
   
  stat_smooth(method="glm", method.args = list(family = "binomial"))+
  facet_wrap(sector~.)+theme_bw()+
               xlab("Maximum temperature (°C)")+
              ylab("Minimum fractional sea ice cover")
              
  ggplot(subset(adfm,Season=="AMJ"|Season=="JAS"),aes(Year,tMax))+
  geom_point(alpha=0.05)+
  geom_smooth(method="gam",formula=y~s(x,k=8))+
  theme_bw()+facet_wrap(sector~.)+
  ylab("Maximum temperature (°C)")+

    
# figure 5    
ggplot(subset(adfm,Season=="AMJ"|Season=="JAS"),aes(Year,sicMin))+
  geom_point(alpha=0.05)+
  geom_smooth(method="gam",formula=y~s(x,k=8))+
  theme_bw()+facet_wrap(sector~.)+
  ylab("Minimum fractional sea ice cover")


```

### krill length data from fishing

```{r}

### reprocessing climate data 
### we know sea  might be important during winter, and temperature influences summer productivity

# first, lets re-evaluate events date based on the peak date

evenT$Month<-month(evenT$date_peak)
evenS$Month<-month(evenS$date_peak)

evenT$Year<-year(evenT$date_peak)
evenT$quart<-quarter(evenT$date_peak) # tri-mensual classification

evenS$Year<-year(evenS$date_peak)
evenS$quart<-quarter(evenS$date_peak) # tri-mensual classification


# change names
evenT$Season[evenT$quart=="1"]<-"JFM"
evenT$Season[evenT$quart=="2"]<-"AMJ"
evenT$Season[evenT$quart=="3"]<-"JAS"
evenT$Season[evenT$quart=="4"]<-"OND"

evenS$Season[evenS$quart=="1"]<-"JFM"
evenS$Season[evenS$quart=="2"]<-"AMJ"
evenS$Season[evenS$quart=="3"]<-"JAS"
evenS$Season[evenS$quart=="4"]<-"OND"



WseaIce<-plyr::ddply(subset(evenS,Season=="AMJ"|Season=="JAS"), c("Year","sector"), summarise,
                     TN=length(duration),
                     TmeanDur=mean(duration),
                     TcumDur=sum(duration),
                     TmeanInt=mean(intensity_mean),
                     TmaxInt=max(intensity_max))

Stemp<-plyr::ddply(subset(evenS,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                     SN=length(duration),
                     SmeanDur=mean(duration),
                     ScumDur=sum(duration),
                     SmeanInt=mean(intensity_mean),
                     SmaxInt=max(intensity_max))


evesum<-merge(WseaIce,Stemp,by=c("Year","sector"),all=T)



# load climatology data 

climS<-readRDS("D:/HeatWaves/Krill/Data/climatologyICE.Rds")
climT<-readRDS("D:/HeatWaves/Krill/Data/climatologyTEMP.Rds")

clim<-merge(climT,climS,by=c("t","sector"))


clim$Year<-year(clim$t)
  clim$Month<-month(clim$t)
  clim$Week<-week(clim$t)
  
  clim$Quarter<-quarter(clim$t)
  
  clim$Season[clim$Quarter=="1"]<-"JFM"
  clim$Season[clim$Quarter=="2"]<-"AMJ"
  clim$Season[clim$Quarter=="3"]<-"JAS"
  clim$Season[clim$Quarter=="4"]<-"OND"
  
  head(clim)
 
  climWS<-plyr::ddply(subset(clim,Season=="AMJ"|Season=="JAS"), c("Year","sector"), summarise,
                    SICPeak=max(sicMean),SICmean=mean(sicMean),thresh.y=mean(thresh.y))

climST<-plyr::ddply(subset(clim,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   
                    TempMax=max(tMax),TempMean=mean(tMax),thresh.x=mean(thresh.x))
climWS<-merge(climWS,climST)


```

### load and process krill data

```{r}

  
krill.strata<-readRDS("KrillData (2).Rdata") 


krill.strata$timestamp<-as.POSIXct(strptime(krill.strata$datetime_set_start, format="%Y-%m-%d %H:%M:%S", tz="GMT"))
krill.strata$Week<-week(krill.strata$timestamp)
krill.strata$Year<-year(krill.strata$timestamp)
krill.strata$Month<-month(krill.strata$timestamp)
krill.strata$Quarter<-quarter(krill.strata$timestamp)

krill.strata$Season[krill.strata$Quarter=="1"]<-"JFM"
krill.strata$Season[krill.strata$Quarter=="2"]<-"AMJ"
krill.strata$Season[krill.strata$Quarter=="3"]<-"JAS"
krill.strata$Season[krill.strata$Quarter=="4"]<-"OND"

krill.strata$sector<-krill.strata$ID


krill.strata$stage<-as.factor(krill.strata$maturity_stage)

krist<-na.omit(data.frame(krill.strata[13],krill.strata[,16:23]))

#correct strata name to match the climatology data

krist$sector[krist$sector=="Extra"]<-"EX"
krist$sector[krist$sector=="SSIW"]<-"SS"

krist$Talla<-ifelse(krist$length_total_cm<4,"Recruit","Adult") # first step to calculate the recruitment index


ad<-(plyr::ddply(krist, c("Year","sector"), summarise,
                        Ntot=length(na.omit(length_total_cm))))  # number of adults

rec<-(plyr::ddply(subset(krist,Talla=="Recruit"), c("Year","sector"), summarise,
                        Nrec=length(na.omit(length_total_cm))))  # number of recruiters


recind<-merge(ad,rec,all.x=TRUE)  # merge to calculate recruitment index

recind$Nrec[is.na(recind$Nrec)]<-0


recind$recind<-recind$Nrec/recind$Ntot  # recruitment index

summary(recind$Year)


```

#First question: does recruitment is related to sea ice and temperature variability?

```{r}

cwsk<-merge(recind,climWS)


ggplot(cwsk,aes(SICmean,recind))+geom_point()+
  
  theme_bw()+
  stat_smooth(method="glm", method.args = list(family = "binomial"))+
  xlab("Proportion")+ylab("Recruitment index")+
  ggtitle(label="a. Mean winter sea ice cover")+
  
  ggplot(cwsk,aes(TempMax,recind))+geom_point()+
  
  theme_bw()+
  stat_smooth(method="glm", method.args = list(family = "binomial"))+
  xlab("°C")+ylab("Recruitment index")+
  ggtitle(label="b. Mean maximum summer temperature")

lmp1<-lmPerm::lmp(recind~SICmean*TempMax,data=cwsk,perm="Prob",center=T,singular.ok = F)
summary(lmp1)
anova(lmp1)
```
#second question: does extreme events of low sea ice and high sea temperature affect recruitment?
```{r}
eveK<-merge(recind,evesum,all=T)
summary(eveK)

eveK$TN[is.na(eveK$TN)]<-0
eveK$SN[is.na(eveK$SN)]<-0


eveK$TempCases<-ifelse(eveK$TN>0,"One or more","No cases")
eveK$IceCases<-ifelse(eveK$SN>0,"One or more","No cases")

eveK2<-na.omit(data.frame(eveK[1:2],eveK[5],eveK[16:17]))

ggplot(eveK2,aes(IceCases,recind))+geom_boxplot()+

  theme_bw()+
  
  ggplot(eveK2,aes(TempCases,recind))+geom_boxplot()+

  theme_bw()


lmp3<-lmPerm::lmp(recind~IceCases*TempCases,data=eveK2,perm="Prob",center=T,singular.ok = F)

summary(lmp3)

anova(lmp3)

```


### conclusion: winter sea ice cover explais better the year-round recruitment...

#### now that we have this information, let's see whether chlorophyll relates to sea ice 

```{r,echo=T}

# chlorohyll data


#rerddap::info(datasetid = "erdMH1chlamday", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

###
dfL<-rbind(df1,df2,df3,df4,df9)


  OISST_CHL<- rerddap::griddap(datasetx = "erdMH1chlamday",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c("2015-01-01T12:00:00Z", "2022-03-31T12:00:00Z"), 
                                #zlev = c(0, 0),
                                latitude = c(min(dfL$ymin),max(dfL$ymax)), 
                                longitude = c(min(dfL$xmin),max(dfL$xmax)),
                                fields = c("chlorophyll"))$data %>%
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, chl = chlorophyll, lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, chl) %>% 
    stats::na.omit()

base::saveRDS(OISST_CHL, file = "D:/HeatWaves/Krill/Data/OISST_CHL.Rds")


CHL<-OISST_CHL

CHL$Year<-year(CHL$t)
CHL$Month<-month(CHL$t)


CHLm<-plyr::ddply(CHL, c("Year","Month"), summarise,
                  chlm=mean(chl),chlse=sd(chl)/sqrt(length(chl)-1))

summary(CHLm$Year)

SS<-readRDS("D:/HeatWaves/Krill/Data/OISST_SS3.Rds")
EI<-readRDS("D:/HeatWaves/Krill/Data/OISST_EL3.Rds")
BS<-readRDS("D:/HeatWaves/Krill/Data/OISST_BS3.Rds")
EX<-readRDS("D:/HeatWaves/Krill/Data/OISST_EX3.Rds")
JI<-readRDS("D:/HeatWaves/Krill/Data/OISST_JI3.Rds")


SS$Year<-year(SS$t)
SS$Month<-month(SS$t)

EI$Year<-year(EI$t)
EI$Month<-month(EI$t)

BS$Year<-year(BS$t)
BS$Month<-month(BS$t)

EX$Year<-year(EX$t)
EX$Month<-month(EX$t)

JI$Year<-year(JI$t)
JI$Month<-month(JI$t)

SSm<-plyr::ddply(SS, c("Year","Month"), summarise,sicm=mean(sic),sicsd=sd(sic),sstm=mean(temp),sstsd=sd(temp))
EIm<-plyr::ddply(EI, c("Year","Month"), summarise,sicm=mean(sic),sicsd=sd(sic),sstm=mean(temp),sstsd=sd(temp))
BSm<-plyr::ddply(BS, c("Year","Month"), summarise,sicm=mean(sic),sicsd=sd(sic),sstm=mean(temp),sstsd=sd(temp))
EXm<-plyr::ddply(EX, c("Year","Month"), summarise,sicm=mean(sic),sicsd=sd(sic),sstm=mean(temp),sstsd=sd(temp))
JIm<-plyr::ddply(JI, c("Year","Month"), summarise,sicm=mean(sic),sicsd=sd(sic),sstm=mean(temp),sstsd=sd(temp))


all<-rbind(SSm,EIm,BSm,EXm,JIm)

allm<-plyr::ddply(all, c("Year","Month"), summarise,sic=mean(sicm),sicsd=mean(sicsd),sst=mean(sstm),sstsd=mean(sstsd))

chl.df<-merge(allm,CHLm,all.x=T)

ggplot(chl.df,aes(Month,sic))+geom_point()


library(lubridate)
library(tidyr)
library(tidyquant)
library(dplyr)
library(broom)
library(purrr)
library(stringr)
library(knitr)
library(timetk)


chl.df$chlm[is.na(chl.df$chlm) & chl.df$Year=="2015"]<-min(na.omit(chl.df$chlm[chl.df$Year=="2015"]))
chl.df$chlm[is.na(chl.df$chlm) & chl.df$Year=="2016"]<-min(na.omit(chl.df$chlm[chl.df$Year=="2016"]))
chl.df$chlm[is.na(chl.df$chlm) & chl.df$Year=="2017"]<-min(na.omit(chl.df$chlm[chl.df$Year=="2017"]))
chl.df$chlm[is.na(chl.df$chlm) & chl.df$Year=="2018"]<-min(na.omit(chl.df$chlm[chl.df$Year=="2018"]))
chl.df$chlm[is.na(chl.df$chlm) & chl.df$Year=="2019"]<-min(na.omit(chl.df$chlm[chl.df$Year=="2019"]))
chl.df$chlm[is.na(chl.df$chlm) & chl.df$Year=="2020"]<-min(na.omit(chl.df$chlm[chl.df$Year=="2020"]))
chl.df$chlm[is.na(chl.df$chlm) & chl.df$Year=="2021"]<-min(na.omit(chl.df$chlm[chl.df$Year=="2021"]))
#chl.df$chlm[is.na(chl.df$chlm) & chl.df$Year=="2022"]<-min(na.omit(chl.df$chlm[chl.df$Year=="2022"]))


(ggplot(chl.df,aes(as.factor(Month),chlm))+
  geom_boxplot()+ylab("Chlorophyll-a concentration (mg/m3)")+
 theme_bw())/

( ggplot(chl.df,aes(as.factor(Month),sst))+
   geom_boxplot()+ylab("Sea surface temperature (°C)")+
 theme_bw())/
 
( ggplot(chl.df,aes(as.factor(Month),sic))+
   geom_boxplot()+ylab("Sea ice cover (%)")+
 theme_bw())

chl.df$Month<-ifelse(as.numeric(chl.df$Month)<10,paste(0,chl.df$Month,sep=""),chl.df$Month)
head(chl.df)


chl.df$ym<-as.POSIXct(strptime(paste(chl.df$Year,chl.df$Month,"01",sep="-"), format="%Y-%m-%d", tz="GMT"))

head(chl.df)


chl.df2<-data.frame(chl.df[1:3],chl.df[5],chl.df[7],chl.df[9])

chlts<-xts(chl.df2,order.by=chl.df2$ym)

head(chlts)

kel<-data.frame(chlts[,1:6],chlts$chlm %>%           
                  lag.xts(k = 1:6))
head(kel)

names(kel)[names(kel) == 'chlm.6'] <- 'Lag1'
names(kel)[names(kel) == 'chlm.1'] <- 'Lag2'
names(kel)[names(kel) == 'chlm.2'] <- 'Lag3'
names(kel)[names(kel) == 'chlm.3'] <- 'Lag4'
names(kel)[names(kel) == 'chlm.4'] <- 'Lag5'
names(kel)[names(kel) == 'chlm.5'] <- 'Lag6'
names(kel)[names(kel) == 'chlm'] <- 'Lag0'
head(kel)

kelm<-melt(kel,id.vars=c("Year","Month","sic","sst","ym"))

kelm$LagN<-as.numeric(substring(kelm$variable,first=4,last=4))
head(kelm)

names(kelm)[names(kelm) == 'variable'] <- 'Lag'
names(kelm)[names(kelm) == 'value'] <- 'chlm'

kelm$chlm<-as.numeric(kelm$chlm)
kelm$sic<-as.numeric(kelm$sic)
kelm$sst<-as.numeric(kelm$sst)


head(kelm)

kelm$quarter<-quarter(kelm$ym)
kelm$Season[kelm$quarter=="1"]<-"JFM"
kelm$Season[kelm$quarter=="2"]<-"AMJ"
kelm$Season[kelm$quarter=="3"]<-"JAS"
kelm$Season[kelm$quarter=="4"]<-"OND"


kelm$Season<-factor(kelm$Season,levels=c("JFM","AMJ","JAS","OND"))


(ggplot(subset(kelm,LagN>0 & Season=="JFM"),aes(sic,chlm))+
  geom_smooth(method="gam")+theme_bw()+
  geom_point()+facet_wrap(Lag~.))+ggtitle(label="a. JFM")+
  xlab("Sea ice cover (%)")+ylab("Chlorophyll-a concentration (mg/m3)")+

(ggplot(subset(kelm,LagN>0 & Season=="AMJ"),aes(sic,chlm))+
    geom_smooth(method="gam")+theme_bw()+
    geom_point()+facet_wrap(Lag~.))+ggtitle(label="b. AMJ")+
  xlab("Sea ice cover (%)")+ylab("Chlorophyll-a concentration (mg/m3)")+

(ggplot(subset(kelm,LagN>0 & Season=="JAS"),aes(sic,chlm))+
    geom_smooth(method="gam")+theme_bw()+
    geom_point()+facet_wrap(Lag~.))+ggtitle(label="c. JAS")+
  xlab("Sea ice cover (%)")+ylab("Chlorophyll-a concentration (mg/m3)")+

(ggplot(subset(kelm,LagN>0 & Season=="OND"),aes(sic,chlm))+
    geom_smooth(method="gam")+theme_bw()+
    geom_point()+facet_wrap(Lag~.))+ggtitle(label="d. OND")+
  xlab("Sea ice cover (%)")+ylab("Chlorophyll-a concentration (mg/m3)")

(ggplot(subset(kelm,LagN>0 & Season=="JFM"),aes(sst,chlm))+
  geom_smooth(method="gam")+theme_bw()+
  geom_point()+facet_wrap(Lag~.))+ggtitle(label="a. JFM")+
  xlab("Sea surface temperature (°C)")+ylab("Chlorophyll-a concentration (mg/m3)")+
  
  (ggplot(subset(kelm,LagN>0 & Season=="AMJ"),aes(sst,chlm))+
     geom_smooth(method="gam")+theme_bw()+
     geom_point()+facet_wrap(Lag~.))+ggtitle(label="b. AMJ")+
  xlab("Sea surface temperature (°C)")+ylab("Chlorophyll-a concentration (mg/m3)")+
  
  (ggplot(subset(kelm,LagN>0 & Season=="JAS"),aes(sst,chlm))+
     geom_smooth(method="gam")+theme_bw()+
     geom_point()+facet_wrap(Lag~.))+ggtitle(label="c. JAS")+
  xlab("Sea surface temperature (°C)")+ylab("Chlorophyll-a concentration (mg/m3)")+
  
  (ggplot(subset(kelm,LagN>0 & Season=="OND"),aes(sst,chlm))+
     geom_smooth(method="gam")+theme_bw()+
     geom_point()+facet_wrap(Lag~.))+ggtitle(label="d. OND")+
  xlab("Sea surface temperature (°C)")+ylab("Chlorophyll-a concentration (mg/m3)")




```

# Sea ice during Autumn (AMJ) affects chlorophyll concentration with a 6-months delay


```{r}



wsic<-subset(kelm,Month=="06"|Month=="07"|Month=="08"|Month=="09")
wsicL<-subset(wsic,LagN>3)

summary(wsicL$sic[wsicL$Month=="06"])
summary(wsicL$sic[wsicL$Month=="07"])
summary(wsicL$sic[wsicL$Month=="08"])
summary(wsicL$sic[wsicL$Month=="09"])

wsicL$catsic<-ifelse(wsicL$sic<0.47 & wsicL$Month=="06","Low","High")
wsicL$catsic<-ifelse(wsicL$sic<0.51 & wsicL$Month=="07","Low","High")
wsicL$catsic<-ifelse(wsicL$sic<0.54 & wsicL$Month=="08","Low","High")
wsicL$catsic<-ifelse(wsicL$sic<0.45 & wsicL$Month=="09","Low","High")


ggplot(wsicL,aes(catsic,chlm))+geom_boxplot()
```

#in winters when sea ice was low during autumn/winter, there was less chlorophyl between 4 to 6 months later




```{r}


ggplot(wsicL,aes(sst,sic))+geom_point()



```


#winter higher temperatures meant lower winter sea ice. Do increased temperature in summer compensates for this?


```{r}


wsic<-subset(kelm,Month=="06"|Month=="07"|Month=="08"|Month=="09")
wsicL<-subset(wsic,LagN>3)

summary(wsicL$sic[wsicL$Month=="06"])
summary(wsicL$sic[wsicL$Month=="07"])
summary(wsicL$sic[wsicL$Month=="08"])
summary(wsicL$sic[wsicL$Month=="09"])

wsicL$catsic<-ifelse(wsicL$sic<0.47 & wsicL$Month=="06","Low","High")
wsicL$catsic<-ifelse(wsicL$sic<0.51 & wsicL$Month=="07","Low","High")
wsicL$catsic<-ifelse(wsicL$sic<0.54 & wsicL$Month=="08","Low","High")
wsicL$catsic<-ifelse(wsicL$sic<0.45 & wsicL$Month=="09","Low","High")


ggplot(wsicL,aes(catsic,chlm))+geom_boxplot()

ggplot(wsicL,aes(sst,sic))+geom_point()



sumt<-subset(kelm,Month=="12"|Month=="01"|Month=="02"|Month=="03")
sumtL<-subset(sumt,LagN<4)

summary(sumtL$sst[sumtL$Month=="12"])
summary(sumtL$sst[sumtL$Month=="01"])
summary(sumtL$sst[sumtL$Month=="02"])
summary(sumtL$sst[sumtL$Month=="03"])

sumtL$cattemp<-ifelse(sumtL$sst<(-0.47) & sumtL$Month=="12","Low","High")
sumtL$cattemp<-ifelse(sumtL$sst<(-0.04) & sumtL$Month=="01","Low","High")
sumtL$cattemp<-ifelse(sumtL$sst<(0.45) & sumtL$Month=="02","Low","High")
sumtL$cattemp<-ifelse(sumtL$sst<(0.026) & sumtL$Month=="03","Low","High")


ggplot(wsicL,aes(catsic,chlm))+geom_boxplot()+theme_bw()+
  xlab("Sea ice cover during winter")+ylab("Chlorophyll-a concentration in summer")+
  ggtitle("a.")+
ggplot(sumtL,aes(cattemp,chlm))+geom_boxplot()+theme_bw()+
  xlab("Sea surface temperature during summer")+ylab("Chlorophyll-a concentration in summer")+
  ggtitle("b.")

```


# my answer is no, higher summer temperatures do not compensate for the winter loss of ice.