---
title: Detection of extreme heatwaves and low sea ice cover in the Antarctic Peninsula  for
  evaluation of effects over Krill, Part I - download data
author: "Lucas Krüger, Maurício Mardones, Lorena Rebolledo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#---------- Download NOAA erddap optimum interpolation SST----

#this part of the script is a modification of the scripts in the heatwaveR vignette
#available at https://robwschlegel.github.io/heatwaveR/articles/OISST_preparation.html


```{r, echo =FALSE}
#---------- Download NOAA erddap optimum interpolation SST----

# The packages we will use
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(patchwork)
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing
library(raster)
library(terra)
library(sf)

library(mapdata)
library(mapproj)

library(reshape2)

```

# The information for the NOAA OISST data

```{r, echo=FALSE}


rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

rerddap::info(datasetid = "erdMH1chlamday", url = "https://coastwatch.pfeg.noaa.gov/erddap/")



```

# Download and prepare data based on user provided start and end dates using three different requests per geographical strata so it can be run on slow PCs.

```{r}


# spatial units to use


shp<-shapefile("Clipped_StrataGCS.shp")


crs=CRS( "+proj=longlat +datum=WGS84 +no_defs") # coordinate reference system CRS

shp2<-spTransform(shp,crs) # be sure it is in the correct coordinate system


#load antarctic peninsula shapefile to plot it
# this shapefile was cut from high resolution file from the SCAR Antarctic Digital Database, 2023
# Gerrish, L., Ireland, L., Fretwell, P., & Cooper, P. (2023). High resolution vector polygons of the Antarctic coastline (7.7) [Data set]. UK Polar Data Centre, Natural Environment Research Council, UK Research & Innovation. https://doi.org/10.5285/0be5339c-9d35-44c9-a10f-da4b5356840b'


shp.sf<-st_as_sf(shp)
unique(shp.sf$ID)

#Extract coordinates and generates a bounding box 

geom<-st_geometry(shp.sf)

bbox_list <- lapply(st_geometry(shp.sf), st_bbox)

#min and max coordinates of each strata as data frame

df1<-data.frame(t(as.data.frame((bbox_list[[1]])))) # Elephant
df2<-data.frame(t(as.data.frame((bbox_list[[2]])))) # South Shetlands
df3<-data.frame(t(as.data.frame((bbox_list[[3]])))) # Bransfield
df4<-data.frame(t(as.data.frame((bbox_list[[4]])))) #Joinville
df5<-data.frame(t(as.data.frame((bbox_list[[5]])))) # Outer 
df9<-data.frame(t(as.data.frame((bbox_list[[9]])))) #Extra SW

```
```{r,echo=FALSE}

library(heatwaveR)


#load and merge data

BS<-base::readRDS("OISST_BS1.Rds")
EI<-base::readRDS("OISST_EL1.Rds")
EX<-base::readRDS("OISST_EX1.Rds")
JI<-base::readRDS("OISST_JI1.Rds")
SS<-base::readRDS("OISST_SS1.Rds")
SW<-base::readRDS("OISST_SW1.Rds")


```


### download chlorophyll-a data

```{r}

# chlorohyll data


rerddap::info(datasetid = "erdMH1chlamday", url = "https://coastwatch.pfeg.noaa.gov/erddap/")


dfL<-rbind(df1,df2,df3,df4,df9)


OISST_CHL<- rerddap::griddap(datasetx = "erdMH1chlamday",
                              url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                              time = c("2003-01-16T12:00:00Z", "2022-05-16T00:00:00Z"), 
                             #zlev = c(0, 0),
                            latitude = c(min(dfL$ymin),max(dfL$ymax)), 
                             longitude = c(min(dfL$xmin),max(dfL$xmax)),
                              fields = c("chlorophyll"))$data %>%
  dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
  dplyr::rename(t = time, chl = chlorophyll, lon = longitude, lat = latitude) %>% 
  dplyr::select(lon, lat, t, chl) %>% 
  stats::na.omit()

CHL<-OISST_CHL



CHL$Year<-year(CHL$t)
CHL$Month<-month(CHL$t)


CHLm<-plyr::ddply(CHL, c("Year","Month"), summarise,
                  chlm=mean(chl),chlse=sd(chl)/sqrt(length(chl)-1))

summary(CHLm$Year)

chl.bs<-subset(CHL,lon<max(BS$lon) & lon>min(BS$lon) &
                 lat<max(BS$lat) & lat>min(BS$lat))

chl.ei<-subset(CHL,lon<max(EI$lon) & lon>min(EI$lon) &
                 lat<max(EI$lat) & lat>min(EI$lat))

chl.ex<-subset(CHL,lon<max(EX$lon) & lon>min(EX$lon) &
                 lat<max(EX$lat) & lat>min(EX$lat))

chl.ji<-subset(CHL,lon<max(JI$lon) & lon>min(JI$lon) &
                 lat<max(JI$lat) & lat>min(JI$lat))

chl.ss<-subset(CHL,lon<max(SS$lon) & lon>min(SS$lon) &
                 lat<max(SS$lat) & lat>min(SS$lat))

chl.sw<-subset(CHL,lon<max(SW$lon) & lon>min(SW$lon) &
                 lat<max(SW$lat) & lat>min(SW$lat))


chl.bs$sector<-c("BS")
chl.bs$Year<-year(chl.bs$t)
chl.bs$Month<-month(chl.bs$t)
chl.bs$quarter<-quarter(chl.bs$t)

chl.bs$Season[chl.bs$quarter=="1"]<-"JFM"
chl.bs$Season[chl.bs$quarter=="2"]<-"AMJ"
chl.bs$Season[chl.bs$quarter=="3"]<-"JAS"
chl.bs$Season[chl.bs$quarter=="4"]<-"OND"

chlbs<-plyr::ddply(subset(chl.bs,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                          chlm=mean(chl),chlsd=sd(chl))


chl.ei$sector<-c("EI")
chl.ei$Year<-year(chl.ei$t)
chl.ei$Month<-month(chl.ei$t)
chl.ei$quarter<-quarter(chl.ei$t)

chl.ei$Season[chl.ei$quarter=="1"]<-"JFM"
chl.ei$Season[chl.ei$quarter=="2"]<-"AMJ"
chl.ei$Season[chl.ei$quarter=="3"]<-"JAS"
chl.ei$Season[chl.ei$quarter=="4"]<-"OND"

chlei<-plyr::ddply(subset(chl.ei,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))



chl.ex$sector<-c("EX")
chl.ex$Year<-year(chl.ex$t)
chl.ex$Month<-month(chl.ex$t)
chl.ex$quarter<-quarter(chl.ex$t)

chl.ex$Season[chl.ex$quarter=="1"]<-"JFM"
chl.ex$Season[chl.ex$quarter=="2"]<-"AMJ"
chl.ex$Season[chl.ex$quarter=="3"]<-"JAS"
chl.ex$Season[chl.ex$quarter=="4"]<-"OND"

chlex<-plyr::ddply(subset(chl.ex,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))



chl.ji$sector<-c("JI")
chl.ji$Year<-year(chl.ji$t)
chl.ji$Month<-month(chl.ji$t)
chl.ji$quarter<-quarter(chl.ji$t)

chl.ji$Season[chl.ji$quarter=="1"]<-"JFM"
chl.ji$Season[chl.ji$quarter=="2"]<-"AMJ"
chl.ji$Season[chl.ji$quarter=="3"]<-"JAS"
chl.ji$Season[chl.ji$quarter=="4"]<-"OND"

chlji<-plyr::ddply(subset(chl.ji,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))



chl.ss$sector<-c("SS")
chl.ss$Year<-year(chl.ss$t)
chl.ss$Month<-month(chl.ss$t)
chl.ss$quarter<-quarter(chl.ss$t)

chl.ss$Season[chl.ss$quarter=="1"]<-"JFM"
chl.ss$Season[chl.ss$quarter=="2"]<-"AMJ"
chl.ss$Season[chl.ss$quarter=="3"]<-"JAS"
chl.ss$Season[chl.ss$quarter=="4"]<-"OND"

chlss<-plyr::ddply(subset(chl.ss,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))



chl.sw$sector<-c("SW")
chl.sw$Year<-year(chl.sw$t)
chl.sw$Month<-month(chl.sw$t)
chl.sw$quarter<-quarter(chl.sw$t)

chl.sw$Season[chl.sw$quarter=="1"]<-"JFM"
chl.sw$Season[chl.sw$quarter=="2"]<-"AMJ"
chl.sw$Season[chl.sw$quarter=="3"]<-"JAS"
chl.sw$Season[chl.sw$quarter=="4"]<-"OND"

chlsw<-plyr::ddply(subset(chl.sw,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))


chlall<-rbind(chlbs,chlei,chlex,chlji,chlss,chlsw)  # summer chlorophyll 

head(chlall)


base::saveRDS(chlall, file = "CHL_summer.Rds")



```

