---
title: Detection of extreme heatwaves and low sea ice cover in the Antarctic Peninsula  for
  evaluation of effects over Krill, Part I - download data
author: "Lucas Krüger, Maurício Mardones, Lorena Rebolledo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#---------- Download NOAA erddap optimum interpolation SST----

#this part of the script is a modification of the scripts in the heatwaveR vignette
#available at https://robwschlegel.github.io/heatwaveR/articles/OISST_preparation.html


```{r, echo =FALSE}
#---------- Download NOAA erddap optimum interpolation SST----

# The packages we will use
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(patchwork)
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing
library(raster)
library(terra)
library(sf)

library(mapdata)
library(mapproj)

library(reshape2)

```

# The information for the NOAA OISST data

```{r, echo=FALSE}


rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

rerddap::info(datasetid = "erdMH1chlamday", url = "https://coastwatch.pfeg.noaa.gov/erddap/")



```

# Download and prepare data based on user provided start and end dates using three different requests per geographical strata so it can be run on slow PCs.

```{r}


# spatial units to use


shp<-shapefile("D:/HeatWaves/Krill/Data/Shapefiles/Clipped_StrataGCS.shp")


crs=CRS( "+proj=longlat +datum=WGS84 +no_defs") # coordinate reference system CRS

shp2<-spTransform(shp,crs) # be sure it is in the correct coordinate system


#load antarctic peninsula shapefile to plot it
# this shapefile was cut from high resolution file from the SCAR Antarctic Digital Database, 2023
# Gerrish, L., Ireland, L., Fretwell, P., & Cooper, P. (2023). High resolution vector polygons of the Antarctic coastline (7.7) [Data set]. UK Polar Data Centre, Natural Environment Research Council, UK Research & Innovation. https://doi.org/10.5285/0be5339c-9d35-44c9-a10f-da4b5356840b'

wap<-shapefile("D:/HeatWaves/Krill/Data/Shapefiles/peninsula_correct.shp") 


# plot map with shapefile (Fig 1)
shpdf<-data.frame(shp2)

ggplot()+
  
  xlab("Longitude")+ylab("Latitude")+
  theme_bw()+
    geom_polygon(data=shp2,aes(x = long, y = lat, group = group,label=group),fill="lightblue",colour="black")+
  geom_polygon(data=wap,aes(x = long, y = lat, group = group))+
geom_text(data=shpdf,aes(label=ID, 
            x=Labx,
            y=Laby),colour="red2",
            size = 3, hjust=0, vjust=-1) +
      coord_quickmap(xlim=c(-75,-50),ylim=c(-75, -59), expand = TRUE, clip = "on")


shp.sf<-st_as_sf(shp)
unique(shp.sf$ID)

#Extract coordinates and generates a bounding box 

geom<-st_geometry(shp.sf)

bbox_list <- lapply(st_geometry(shp.sf), st_bbox)

#min and max coordinates of each strata as data frame

df1<-data.frame(t(as.data.frame((bbox_list[[1]])))) # Elephant
df2<-data.frame(t(as.data.frame((bbox_list[[2]])))) # South Shetlands
df3<-data.frame(t(as.data.frame((bbox_list[[3]])))) # Bransfield
df4<-data.frame(t(as.data.frame((bbox_list[[4]])))) #Joinville
df5<-data.frame(t(as.data.frame((bbox_list[[5]])))) # Outer 
df9<-data.frame(t(as.data.frame((bbox_list[[9]])))) #Extra SW

```

#download and process data for Elephant Island Strata

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df1$ymin),max(df1$ymax)), 
                                longitude = c(min(df1$xmin),max(df1$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_EL1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_EL2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_EL3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_EL1, file = "D:/HeatWaves/Krill/Data/OISST_EL1.Rds")
base::saveRDS(OISST_EL2, file = "D:/HeatWaves/Krill/Data/OISST_EL2.Rds")
base::saveRDS(OISST_EL3, file = "D:/HeatWaves/Krill/Data/OISST_EL3.Rds")

### summarize data per day

EL1<-plyr::ddply(OISST_EL1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EL2<-plyr::ddply(OISST_EL2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EL3<-plyr::ddply(OISST_EL3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

EL12<-rbind(EL1,EL2)
EL123<-rbind(EL12,EL3)

base::saveRDS(EL123, file = "D:/HeatWaves/Krill/Data/Elephant_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_EL1","OISST_EL2","OISST_EL3","OISST_sub_dl","shp","maxmin","bbox_list",
          "EL1","EL2","EL3","EL12","EL123"))

gc()



```
#download and process data for South Shetland West Strata

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df2$ymin),max(df2$ymax)),
                                longitude = c(min(df2$xmin),max(df2$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_SS1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_SS2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_SS3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_SS1, file = "D:/HeatWaves/Krill/Data/OISST_SS1.Rds")
base::saveRDS(OISST_SS2, file = "D:/HeatWaves/Krill/Data/OISST_SS2.Rds")
base::saveRDS(OISST_SS3, file = "D:/HeatWaves/Krill/Data/OISST_SS3.Rds")

### summarize data per day

SS1<-plyr::ddply(OISST_SS1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SS2<-plyr::ddply(OISST_SS2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SS3<-plyr::ddply(OISST_SS3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

SS12<-rbind(SS1,SS2)
SS123<-rbind(SS12,SS3)

base::saveRDS(SS123, file = "D:/HeatWaves/Krill/Data/West_SSI_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_SS1","OISST_SS2","OISST_SS3","OISST_sub_dl",
          "SS1","SS2","SS3","SS12","SS123"))

gc()



```


#download and process data for Bransfield Strait Strata

```{r, echo=FALSE}

df3$ID[3]

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df3$ymin),max(df3$ymax)),
                                longitude = c(min(df3$xmin),max(df3$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_BS1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_BS2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_BS3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_BS1, file = "D:/HeatWaves/Krill/Data/OISST_BS1.Rds")
base::saveRDS(OISST_BS2, file = "D:/HeatWaves/Krill/Data/OISST_BS2.Rds")
base::saveRDS(OISST_BS3, file = "D:/HeatWaves/Krill/Data/OISST_BS3.Rds")


### summarize data per day

BS1<-plyr::ddply(OISST_BS1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

BS2<-plyr::ddply(OISST_BS2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

BS3<-plyr::ddply(OISST_BS3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

BS12<-rbind(BS1,BS2)
BS123<-rbind(BS12,BS3)

base::saveRDS(BS123, file = "D:/HeatWaves/Krill/Data/Bransfield_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_BS1","OISST_BS2","OISST_BS3","OISST_sub_dl",
          "BS1","BS2","BS3","BS12","BS123"))

gc()



```




#download and process data for Joinville Island Strata

```{r, echo=FALSE}



OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df4$ymin),max(df4$ymax)),
                                longitude = c(min(df4$xmin),max(df4$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_JI1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_JI2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_JI3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_JI1, file = "D:/HeatWaves/Krill/Data/OISST_JI1.Rds")
base::saveRDS(OISST_JI2, file = "D:/HeatWaves/Krill/Data/OISST_JI2.Rds")
base::saveRDS(OISST_JI3, file = "D:/HeatWaves/Krill/Data/OISST_JI3.Rds")


OISST_JI1<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI1.Rds")
OISST_JI2<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI2.Rds")
OISST_JI3<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI3.Rds")


### summarize data per day

JI1<-plyr::ddply(OISST_JI1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

JI2<-plyr::ddply(OISST_JI2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

JI3<-plyr::ddply(OISST_JI3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

JI12<-rbind(JI1,JI2)
JI123<-rbind(JI12,JI3)

base::saveRDS(JI123, file = "D:/HeatWaves/Krill/Data/Joinvile_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_JI1","OISST_JI2","OISST_JI3","OISST_sub_dl",
          "JI1","JI2","JI3","JI12","JI123"))

gc()



```


#download and process data for Extra Strata (Gerlache?)

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df5$ymin),max(df5$ymax)),
                                longitude = c(min(df5$xmin),max(df5$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_EX1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_EX2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_EX3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_EX1, file = "D:/HeatWaves/Krill/Data/OISST_EX1.Rds")
base::saveRDS(OISST_EX2, file = "D:/HeatWaves/Krill/Data/OISST_EX2.Rds")
base::saveRDS(OISST_EX3, file = "D:/HeatWaves/Krill/Data/OISST_EX3.Rds")

### summarize data per day

EX1<-plyr::ddply(OISST_EX1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EX2<-plyr::ddply(OISST_EX2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EX3<-plyr::ddply(OISST_EX3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

EX12<-rbind(EX1,EX2)
EX123<-rbind(EX12,EX3)

base::saveRDS(EX123, file = "D:/HeatWaves/Krill/Data/Extra_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_EX1","OISST_EX2","OISST_EX3","OISST_sub_dl",
          "EX1","EX2","EX3","EX12","EX123"))

gc()



```


#download and process data for Outer SW Strata (Gerlache?)

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df9$ymin),max(df9$ymax)),
                                longitude = c(min(df9$xmin),max(df9$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_SW1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_SW2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_SW3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_SW1, file = "D:/HeatWaves/Krill/Data/OISST_SW1.Rds")
base::saveRDS(OISST_SW2, file = "D:/HeatWaves/Krill/Data/OISST_SW2.Rds")
base::saveRDS(OISST_SW3, file = "D:/HeatWaves/Krill/Data/OISST_SW3.Rds")

### summarize data per day

SW1<-plyr::ddply(OISST_SW1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SW2<-plyr::ddply(OISST_SW2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SW3<-plyr::ddply(OISST_SW3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

SW12<-rbind(SW1,SW2)
SW123<-rbind(SW12,SW3)

base::saveRDS(SW123, file = "D:/HeatWaves/Krill/Data/OuterSW_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_SW1","OISST_SW2","OISST_SW3","OISST_sub_dl",
          "SW1","SW2","SW3","SW12","SW123"))

gc()



```

### download chlorophyll-a data

```{r}

# chlorohyll data


rerddap::info(datasetid = "erdMH1chlamday", url = "https://coastwatch.pfeg.noaa.gov/erddap/")


dfL<-rbind(df1,df2,df3,df4,df9)


OISST_CHL<- rerddap::griddap(datasetx = "erdMH1chlamday",
                              url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                              time = c("2003-01-16T12:00:00Z", "2022-05-16T00:00:00Z"), 
                             #zlev = c(0, 0),
                            latitude = c(min(dfL$ymin),max(dfL$ymax)), 
                             longitude = c(min(dfL$xmin),max(dfL$xmax)),
                              fields = c("chlorophyll"))$data %>%
  dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
  dplyr::rename(t = time, chl = chlorophyll, lon = longitude, lat = latitude) %>% 
  dplyr::select(lon, lat, t, chl) %>% 
  stats::na.omit()

CHL<-OISST_CHL

head(BS)

CHL$Year<-year(CHL$t)
CHL$Month<-month(CHL$t)


CHLm<-plyr::ddply(CHL, c("Year","Month"), summarise,
                  chlm=mean(chl),chlse=sd(chl)/sqrt(length(chl)-1))

summary(CHLm$Year)

chl.bs<-subset(CHL,lon<max(BS$lon) & lon>min(BS$lon) &
                 lat<max(BS$lat) & lat>min(BS$lat))

chl.ei<-subset(CHL,lon<max(EI$lon) & lon>min(EI$lon) &
                 lat<max(EI$lat) & lat>min(EI$lat))

chl.ex<-subset(CHL,lon<max(EX$lon) & lon>min(EX$lon) &
                 lat<max(EX$lat) & lat>min(EX$lat))

chl.ji<-subset(CHL,lon<max(JI$lon) & lon>min(JI$lon) &
                 lat<max(JI$lat) & lat>min(JI$lat))

chl.ss<-subset(CHL,lon<max(SS$lon) & lon>min(SS$lon) &
                 lat<max(SS$lat) & lat>min(SS$lat))

chl.sw<-subset(CHL,lon<max(SW$lon) & lon>min(SW$lon) &
                 lat<max(SW$lat) & lat>min(SW$lat))


chl.bs$sector<-c("BS")
chl.bs$Year<-year(chl.bs$t)
chl.bs$Month<-month(chl.bs$t)
chl.bs$quarter<-quarter(chl.bs$t)

chl.bs$Season[chl.bs$quarter=="1"]<-"JFM"
chl.bs$Season[chl.bs$quarter=="2"]<-"AMJ"
chl.bs$Season[chl.bs$quarter=="3"]<-"JAS"
chl.bs$Season[chl.bs$quarter=="4"]<-"OND"

chlbs<-plyr::ddply(subset(chl.bs,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                          chlm=mean(chl),chlsd=sd(chl))


chl.ei$sector<-c("EI")
chl.ei$Year<-year(chl.ei$t)
chl.ei$Month<-month(chl.ei$t)
chl.ei$quarter<-quarter(chl.ei$t)

chl.ei$Season[chl.ei$quarter=="1"]<-"JFM"
chl.ei$Season[chl.ei$quarter=="2"]<-"AMJ"
chl.ei$Season[chl.ei$quarter=="3"]<-"JAS"
chl.ei$Season[chl.ei$quarter=="4"]<-"OND"

chlei<-plyr::ddply(subset(chl.ei,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))



chl.ex$sector<-c("EX")
chl.ex$Year<-year(chl.ex$t)
chl.ex$Month<-month(chl.ex$t)
chl.ex$quarter<-quarter(chl.ex$t)

chl.ex$Season[chl.ex$quarter=="1"]<-"JFM"
chl.ex$Season[chl.ex$quarter=="2"]<-"AMJ"
chl.ex$Season[chl.ex$quarter=="3"]<-"JAS"
chl.ex$Season[chl.ex$quarter=="4"]<-"OND"

chlex<-plyr::ddply(subset(chl.ex,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))



chl.ji$sector<-c("JI")
chl.ji$Year<-year(chl.ji$t)
chl.ji$Month<-month(chl.ji$t)
chl.ji$quarter<-quarter(chl.ji$t)

chl.ji$Season[chl.ji$quarter=="1"]<-"JFM"
chl.ji$Season[chl.ji$quarter=="2"]<-"AMJ"
chl.ji$Season[chl.ji$quarter=="3"]<-"JAS"
chl.ji$Season[chl.ji$quarter=="4"]<-"OND"

chlji<-plyr::ddply(subset(chl.ji,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))



chl.ss$sector<-c("SS")
chl.ss$Year<-year(chl.ss$t)
chl.ss$Month<-month(chl.ss$t)
chl.ss$quarter<-quarter(chl.ss$t)

chl.ss$Season[chl.ss$quarter=="1"]<-"JFM"
chl.ss$Season[chl.ss$quarter=="2"]<-"AMJ"
chl.ss$Season[chl.ss$quarter=="3"]<-"JAS"
chl.ss$Season[chl.ss$quarter=="4"]<-"OND"

chlss<-plyr::ddply(subset(chl.ss,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))



chl.sw$sector<-c("SW")
chl.sw$Year<-year(chl.sw$t)
chl.sw$Month<-month(chl.sw$t)
chl.sw$quarter<-quarter(chl.sw$t)

chl.sw$Season[chl.sw$quarter=="1"]<-"JFM"
chl.sw$Season[chl.sw$quarter=="2"]<-"AMJ"
chl.sw$Season[chl.sw$quarter=="3"]<-"JAS"
chl.sw$Season[chl.sw$quarter=="4"]<-"OND"

chlsw<-plyr::ddply(subset(chl.sw,Season=="JFM"|Season=="OND"), c("Year","sector"), summarise,
                   chlm=mean(chl),chlsd=sd(chl))


chlall<-rbind(chlbs,chlei,chlex,chlji,chlss,chlsw)  # summer chlorophyll 

head(chlall)


base::saveRDS(chlall, file = "D:/HeatWaves/Krill/Data/CHL_summer.Rds")



```

