---
title: Detection of extreme heatwaves and low sea ice cover in the Antarctic Peninsula  for
  evaluation of effects over Krill
author: "Lucas Krüger, Maurício Mardones, Lorena Rebolledo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#---------- Download NOAA erddap optimum interpolation SST----

#this part of the script is a modification of the scripts in the heatwaveR vignette
#available at https://robwschlegel.github.io/heatwaveR/articles/OISST_preparation.html


```{r, echo =FALSE}
#---------- Download NOAA erddap optimum interpolation SST----

# The packages we will use
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(patchwork)
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing
library(raster)
library(terra)
library(sf)

```

# The information for the NOAA OISST data

```{r, echo=FALSE}


rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

```

# Download and prepare data based on user provided start and end dates using three different requests per geographical strata so it can be run on slow PCs.

```{r}


# spatial units to use


shp<-shapefile("D:/HeatWaves/Krill/Data/Shapefiles/Clipped_Strata.shp")


crs=CRS( "+proj=longlat +datum=WGS84 +no_defs") # coordinate reference system CRS

shp2<-spTransform(shp,crs)

plot(shp2)

shp.sf<-st_as_sf(shp2)
unique(shp.sf$ID)

#Extract coordinates and generates a bounding box 

geom<-st_geometry(shp.sf)

bbox_list <- lapply(st_geometry(shp.sf), st_bbox)

#min and max coordinates of each strata as data frame

df1<-data.frame(t(as.data.frame((bbox_list[[1]])))) # Elephant
df2<-data.frame(t(as.data.frame((bbox_list[[2]])))) # South Shetlands
df3<-data.frame(t(as.data.frame((bbox_list[[3]])))) # Bransfield
df4<-data.frame(t(as.data.frame((bbox_list[[4]])))) #Joinville
df5<-data.frame(t(as.data.frame((bbox_list[[5]])))) # Outer (gerlache?)
df9<-data.frame(t(as.data.frame((bbox_list[[9]])))) #Extra SW

```

#download and process data for Elephant Island Strata

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df1$ymin),max(df1$ymax)), 
                                longitude = c(min(df1$xmin),max(df1$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_EL1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_EL2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_EL3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_EL1, file = "D:/HeatWaves/Krill/Data/OISST_EL1.Rds")
base::saveRDS(OISST_EL2, file = "D:/HeatWaves/Krill/Data/OISST_EL2.Rds")
base::saveRDS(OISST_EL3, file = "D:/HeatWaves/Krill/Data/OISST_EL3.Rds")

### summarize data per day

EL1<-plyr::ddply(OISST_EL1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EL2<-plyr::ddply(OISST_EL2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EL3<-plyr::ddply(OISST_EL3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

EL12<-rbind(EL1,EL2)
EL123<-rbind(EL12,EL3)

base::saveRDS(EL123, file = "D:/HeatWaves/Krill/Data/Elephant_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_EL1","OISST_EL2","OISST_EL3","OISST_sub_dl","shp","maxmin","bbox_list",
          "EL1","EL2","EL3","EL12","EL123"))

gc()



```
#download and process data for South Shetland West Strata

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df2$ymin),max(df2$ymax)),
                                longitude = c(min(df2$xmin),max(df2$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_SS1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_SS2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_SS3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_SS1, file = "D:/HeatWaves/Krill/Data/OISST_SS1.Rds")
base::saveRDS(OISST_SS2, file = "D:/HeatWaves/Krill/Data/OISST_SS2.Rds")
base::saveRDS(OISST_SS3, file = "D:/HeatWaves/Krill/Data/OISST_SS3.Rds")

### summarize data per day

SS1<-plyr::ddply(OISST_SS1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SS2<-plyr::ddply(OISST_SS2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SS3<-plyr::ddply(OISST_SS3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

SS12<-rbind(SS1,SS2)
SS123<-rbind(SS12,SS3)

base::saveRDS(SS123, file = "D:/HeatWaves/Krill/Data/West_SSI_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_SS1","OISST_SS2","OISST_SS3","OISST_sub_dl",
          "SS1","SS2","SS3","SS12","SS123"))

gc()



```


#download and process data for Bransfield Strait Strata

```{r, echo=FALSE}

df3$ID[3]

OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df3$ymin),max(df3$ymax)),
                                longitude = c(min(df3$xmin),max(df3$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_BS1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_BS2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_BS3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_BS1, file = "D:/HeatWaves/Krill/Data/OISST_BS1.Rds")
base::saveRDS(OISST_BS2, file = "D:/HeatWaves/Krill/Data/OISST_BS2.Rds")
base::saveRDS(OISST_BS3, file = "D:/HeatWaves/Krill/Data/OISST_BS3.Rds")


### summarize data per day

BS1<-plyr::ddply(OISST_BS1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

BS2<-plyr::ddply(OISST_BS2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

BS3<-plyr::ddply(OISST_BS3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

BS12<-rbind(BS1,BS2)
BS123<-rbind(BS12,BS3)

base::saveRDS(BS123, file = "D:/HeatWaves/Krill/Data/Bransfield_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_BS1","OISST_BS2","OISST_BS3","OISST_sub_dl",
          "BS1","BS2","BS3","BS12","BS123"))

gc()



```




#download and process data for Joinville Island Strata

```{r, echo=FALSE}



OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df4$ymin),max(df4$ymax)),
                                longitude = c(min(df4$xmin),max(df4$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_JI1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_JI2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_JI3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_JI1, file = "D:/HeatWaves/Krill/Data/OISST_JI1.Rds")
base::saveRDS(OISST_JI2, file = "D:/HeatWaves/Krill/Data/OISST_JI2.Rds")
base::saveRDS(OISST_JI3, file = "D:/HeatWaves/Krill/Data/OISST_JI3.Rds")


OISST_JI1<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI1.Rds")
OISST_JI2<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI2.Rds")
OISST_JI3<-base::readRDS("D:/HeatWaves/Krill/Data/OISST_JI3.Rds")


### summarize data per day

JI1<-plyr::ddply(OISST_JI1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

JI2<-plyr::ddply(OISST_JI2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

JI3<-plyr::ddply(OISST_JI3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

JI12<-rbind(JI1,JI2)
JI123<-rbind(JI12,JI3)

base::saveRDS(JI123, file = "D:/HeatWaves/Krill/Data/Joinvile_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_JI1","OISST_JI2","OISST_JI3","OISST_sub_dl",
          "JI1","JI2","JI3","JI12","JI123"))

gc()



```


#download and process data for Extra Strata (Gerlache?)

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df5$ymin),max(df5$ymax)),
                                longitude = c(min(df5$xmin),max(df5$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_EX1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_EX2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_EX3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_EX1, file = "D:/HeatWaves/Krill/Data/OISST_EX1.Rds")
base::saveRDS(OISST_EX2, file = "D:/HeatWaves/Krill/Data/OISST_EX2.Rds")
base::saveRDS(OISST_EX3, file = "D:/HeatWaves/Krill/Data/OISST_EX3.Rds")

### summarize data per day

EX1<-plyr::ddply(OISST_EX1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EX2<-plyr::ddply(OISST_EX2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

EX3<-plyr::ddply(OISST_EX3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

EX12<-rbind(EX1,EX2)
EX123<-rbind(EX12,EX3)

base::saveRDS(EX123, file = "D:/HeatWaves/Krill/Data/Extra_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_EX1","OISST_EX2","OISST_EX3","OISST_sub_dl",
          "EX1","EX2","EX3","EX12","EX123"))

gc()



```


#download and process data for Outer SW Strata (Gerlache?)

```{r, echo=FALSE}


OISST_sub_dl <- function(time_df){
  OISST_DATA1 <- rerddap::griddap(datasetx = "ncdcOisst21Agg_LonPM180",
                                url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                                time = c(time_df$start, time_df$end), 
                                zlev = c(0, 0),
                                latitude = c(min(df9$ymin),max(df9$ymax)),
                                longitude = c(min(df9$xmin),max(df9$xmax)),
                                fields = c("sst","ice"))$data %>% 
    dplyr::mutate(time = base::as.Date(stringr::str_remove(time, "T12:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst, sic=ice,lon = longitude, lat = latitude) %>% 
    dplyr::select(lon, lat, t, temp,sic) %>% 
    stats::na.omit()
}




dl_years1 <- data.frame(date_index = 1:3,
                        start = as.Date(c("1985-01-01","1990-01-01","1995-01-01")),
                        
                        end = as.Date(c("1989-12-31","1994-12-31","1999-12-31")))


dl_years2 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2000-01-01", "2005-01-01", "2010-01-01")),
                        
                        end = as.Date(c("2004-12-31", "2009-12-31","2014-12-31")))


dl_years3 <- data.frame(date_index = 1:3,
                        start = as.Date(c("2015-01-01","2020-01-01","2021-01-01")),
                        
                        end = as.Date(c("2019-12-31","2020-12-31","2022-12-31")))


###download and save data


base::system.time(
  OISST_SW1 <- dl_years1 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
) 


base::system.time(
  OISST_SW2 <- dl_years2 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)


base::system.time(
  OISST_SW3 <- dl_years3 %>% 
    dplyr::group_by(date_index) %>% 
    dplyr::group_modify(~OISST_sub_dl(.x)) %>% 
    dplyr::ungroup() %>% 
    dplyr::select(lon, lat, t, temp,sic)
)

base::saveRDS(OISST_SW1, file = "D:/HeatWaves/Krill/Data/OISST_SW1.Rds")
base::saveRDS(OISST_SW2, file = "D:/HeatWaves/Krill/Data/OISST_SW2.Rds")
base::saveRDS(OISST_SW3, file = "D:/HeatWaves/Krill/Data/OISST_SW3.Rds")

### summarize data per day

SW1<-plyr::ddply(OISST_SW1, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SW2<-plyr::ddply(OISST_SW2, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

SW3<-plyr::ddply(OISST_SW3, c("t"), summarise,
            tSD=sd(temp),tMax=max(temp),
            tMin=min(temp),temp=mean(temp),
            sicMean=mean(sic),sicMax=max(sic),
            sicMin=min(sic),sicSD=sd(sic))

### merge data

SW12<-rbind(SW1,SW2)
SW123<-rbind(SW12,SW3)

base::saveRDS(SW123, file = "D:/HeatWaves/Krill/Data/OuterSW_Strata.Rds")

#remove data and release unused memory

rm(list=c("OISST_SW1","OISST_SW2","OISST_SW3","OISST_sub_dl",
          "SW1","SW2","SW3","SW12","SW123"))

gc()



```


###----------- calculating marine heatwaves-------------


```{r}

library(heatwaveR)


#load and merge data

BS<-base::readRDS("D:/HeatWaves/Krill/Data/Bransfield_Strata.Rds")
EI<-base::readRDS("D:/HeatWaves/Krill/Data/Elephant_Strata.Rds")
EX<-base::readRDS("D:/HeatWaves/Krill/Data/Extra_Strata.Rds")
JI<-base::readRDS("D:/HeatWaves/Krill/Data/Joinvile_Strata.Rds")
SS<-base::readRDS("D:/HeatWaves/Krill/Data/West_SSI_Strata.Rds")
SW<-base::readRDS("D:/HeatWaves/Krill/Data/OuterSW_Strata.Rds")

### calculate thresholds from climatology

# tMax thresholds
tMax_clim.BS <- ts2clm(data = BS, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)

tMax_clim.EI <- ts2clm(data = EI, 
                     y = tMax, climatologyPeriod = c("1986-01-01", "2022-12-31"), pctile = 90)

tMax_clim.EX <- ts2clm(data = EX, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)

tMax_clim.JI <- ts2clm(data = JI, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)

tMax_clim.SS <- ts2clm(data = SS, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)

tMax_clim.SW <- ts2clm(data = SW, 
                     y = tMax, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 90)



# iMin thresholds

sicMean_clim.BS <- ts2clm(data = BS, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 25)
sicMean_clim.EI <- ts2clm(data = EI, 
                     y = sicMean, climatologyPeriod = c("1986-01-01", "2022-12-31"), pctile = 25)
sicMean_clim.EX <- ts2clm(data = EX, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 25)

sicMean_clim.JI <- ts2clm(data = JI, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 25)
sicMean_clim.SS <- ts2clm(data = SS, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 25)
sicMean_clim.SW <- ts2clm(data = SW, 
                     y = sicMean, climatologyPeriod = c("1985-01-01", "2022-12-31"), pctile = 25)



#DETECT EVENTS
eBS.t <- detect_event(data = tMax_clim.BS, y = tMax,minDuration = 5,maxGap = 7) 
eEI.t <- detect_event(data = tMax_clim.EI, y = tMax,minDuration = 5,maxGap = 7) 
eEX.t <- detect_event(data = tMax_clim.EX, y = tMax,minDuration = 5,maxGap = 7) 
eJI.t <- detect_event(data = tMax_clim.JI, y = tMax,minDuration = 5,maxGap = 7) 
eSS.t <- detect_event(data = tMax_clim.SS, y = tMax,minDuration = 5,maxGap = 7) 
eSW.t <- detect_event(data = tMax_clim.SW, y = tMax,minDuration = 5,maxGap = 7) 


eBS.sic <- detect_event(data = sicMean_clim.BS, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eEI.sic <- detect_event(data = sicMean_clim.EI, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eEX.sic <- detect_event(data = sicMean_clim.EX, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eJI.sic <- detect_event(data = sicMean_clim.JI, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eSS.sic <- detect_event(data = sicMean_clim.SS, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 
eSW.sic <- detect_event(data = sicMean_clim.SW, y = sicMean,minDuration = 3,maxGap = 7,coldSpells = T) 

bsevenT<-data.frame(eBS.t$event,sector=c("BS")) # Elephant Island events
elevenT<-data.frame(eEI.t$event,sector=c("EI")) # Elephant Island events
exevenT<-data.frame(eEX.t$event,sector=c("EX")) # Elephant Island events
jievenT<-data.frame(eJI.t$event,sector=c("JI")) # Elephant Island events
ssevenT<-data.frame(eSS.t$event,sector=c("SS")) # Elephant Island events
swevenT<-data.frame(eSW.t$event,sector=c("SW")) # Elephant Island events

bsevenS<-data.frame(eBS.sic$event,sector=c("BS")) # Elephant Island events
elevenS<-data.frame(eEI.sic$event,sector=c("EI")) # Elephant Island events
exevenS<-data.frame(eEX.sic$event,sector=c("EX")) # Elephant Island events
jievenS<-data.frame(eJI.sic$event,sector=c("JI")) # Elephant Island events
ssevenS<-data.frame(eSS.sic$event,sector=c("SS")) # Elephant Island events
swevenS<-data.frame(eSW.sic$event,sector=c("SW")) # Elephant Island events

evenT<-rbind(bsevenT,elevenT,exevenT,jievenT,ssevenT,swevenT) # join events data
evenS<-rbind(bsevenS,elevenS,exevenS,jievenS,ssevenS,swevenS) # join events data


evenT$Day<-(decimal_date(evenT$date_peak)-year(evenT$date_peak))*365 # day of the year
evenT$quart<-quarter(evenT$date_peak) # tri-mensual classification

evenS$Day<-(decimal_date(evenS$date_peak)-year(evenS$date_peak))*365 # day of the year
evenS$quart<-quarter(evenS$date_peak) # tri-mensual classification


# change names
evenT$Season[evenT$quart=="1"]<-"JFM"
evenT$Season[evenT$quart=="2"]<-"AMJ"
evenT$Season[evenT$quart=="3"]<-"JAS"
evenT$Season[evenT$quart=="4"]<-"OND"

evenS$Season[evenS$quart=="1"]<-"JFM"
evenS$Season[evenS$quart=="2"]<-"AMJ"
evenS$Season[evenS$quart=="3"]<-"JAS"
evenS$Season[evenS$quart=="4"]<-"OND"

evenT$Season<-factor(evenT$Season,levels=c("JFM","AMJ","JAS","OND")) # order levels

evenS$Season<-factor(evenS$Season,levels=c("JFM","AMJ","JAS","OND")) # order levels

#PLOT

ggplot(evenT,aes(date_peak,duration,colour=Season,linetype=Season))+
 geom_lolli()+
  theme_bw()+ylab("Duration (days)")+
  scale_colour_manual(values=c("red","green4","blue","orange"))+
  facet_wrap(sector~.)+
  xlab("Date peak")

ggplot(evenS,aes(date_peak,duration,colour=Season,linetype=Season))+
 geom_lolli()+
  theme_bw()+ylab("Duration (days)")+
  scale_colour_manual(values=c("red","green4","blue","orange"))+
  facet_wrap(sector~.)+
  xlab("Date peak")


### example for 2022
(event_line(eSW.t, spread = 200, metric = "intensity_cumulative",x=t,y=tMax,
           start_date = "2022-01-01", end_date = "2022-12-31",category=T))/

(event_line(eSW.sic, spread = 200, metric = "intensity_cumulative",x=t,y=sicMean,
           start_date = "2022-01-01", end_date = "2022-12-31",category=T))

```